---
title: "How dates and locations affect the rainfall."
subtitle: "The relationship between dates, locations and rainfall"
author: 
  - Rui Hu
thanks: "Code and data are available at: https://github.com/XYPKQ896/Rainfall_analysis"
date: 09/23/2024
date-format: MM/DD/YYYY
abstract: "Rainfall is associated with citizens. The purpose of this paper is to analyze the rainfall of Toronto. The data on rain gauge locations and rainfall from opendatatoronto contains locations and rainfall. It shows that rainfall is associated with dates and locations."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```


# Introduction

  Rainfall is one of the most important natural phenomena of the environment. It is associated with all humans’ activities. For individuals, we need to decide whether we should take an umbrella this day. For restaurants, how much food they should prepare every morning. A rainy day can’t attract the same number of customers as a sunny day. For farmers, they need to decide the best time to plant or harvest. City authorities should prepare for the possible heavy rain in the future.

  As a result, this paper purposes to understand the rainfall and contribute to the agriculture, environment protection and city planning. We use data from @opendatatoronto in order to disclose the relationship between precipitation, locations and seasons. 
  
\newpage

# Content table
The remainder of this paper is structured as follows. 

@sec-data Data

@sec-model Model

@sec-model_set-up Model set-up

@sec-model_justification Model justification

@sec-results Results

@sec-discussion Discussion

@sec-first-point Random or regular

@sec-second-point Longitude and latitude

@sec-third-point Conclusion

@sec-weeknesses Weaknesses and next steps

@sec-additional Additional data details

@sec-model-details Model details

\newpage

# Data {#sec-data}

This data is sourced from the website The City of Toronto’s Open Data Portal @opendatatoronto.
The data is published by @torontowater

All the variables used can be find at @sec-additional

Simulated data is of daily precipitation(@fig-simulate) which considers that rainfall is randomly every day.

```{r}
#| label: fig-simulate
#| fig-cap: Simulate Precipitation
#| echo: false

simulatedata <- read.csv(here::here("data/raw_data/simulated.csv"))
simulatedata$dates <- as.Date(simulatedata$dates)
ggplot(simulatedata, aes(x = dates, y = rainfall)) +
  geom_line(color = "blue") +
  labs(title = "Simulated Daily Precipitation", 
       x = "Date", 
       y = "Rainfall (mm)")
```

Some of our data is of daily precipitation (@fig-daily), from @opendatatoronto.

```{r}
#| label: fig-daily
#| fig-cap: Daily Precipitation
#| echo: false

daily <- filter(data, data$longitude == -79.4781121, data$latitude == 43.6476803)
ggplot(daily, aes(x = date, y = Total_Rainfall)) +
  geom_line(color = "blue") +
  labs(title = "Daily Precipitation", 
       x = "Date", 
       y = "Rainfall (mm)")
```

In this figure, the location is fixed as (longitude=-79.4781121, latitude=43.6476803). In order to find the rainfall everyday from January to September, daily precipitation shows rainfall in mm.

And also planes (@fig-monthly). Daily precipitation is not able to find a rule for the rainfall. Monthly precipitation gathers elements in daily precipitation.

```{r}
#| label: fig-monthly
#| fig-cap: Monthly Precipitation
#| fig-pos: H
#| echo: false
#| warning: false
#| message: false

monthly <- daily %>% 
  mutate(month = format(date, "%Y-%m")) %>%
  group_by(month) %>%
  summarize(monthly_rainfall = sum(Total_Rainfall))

ggplot(monthly, aes(x = month, y = monthly_rainfall)) +
  geom_bar(stat = "identity", color = "blue") +
  labs(title = "Monthly Precipitation", 
       x = "Month", 
       y = "Rainfall (mm)")


```

Variables in two figures are the following:
  Rainfall: precipitation measured by rain gauge equipment (mm)
  Date: Every day from January to September.
  Month: Every month from January to September.


# Model{#sec-model}

The goal of our modelling strategy is twofold. Firstly, it tells the relationship between precipitation and longitude is positive or negative. Secondly, it tells the relationship between precipitation and latitude is positive or negative.

Here we briefly describe the linear regression model used to investigate the effects of longitude and latitude on precipitation.

More information are included in [Appendix -@sec-model-details].

## Model set-up{#sec-model_set-up}

Define $y$ as the total rainfall in 2024. Then $x_1$ is the longitude and $x_2$ is the latitude.  

\begin{align} 
y = ax_1 +bx_2 + e
\end{align}

We run the model in R [@citeR].


### Model justification{#sec-model_justification}

We expect a positive relationship between the precipitation and coordinates. In particular, increasing longitude or increasing latitude may cause increasing precipitation. $y$ represents precipitation in mm. $x_1$ represents longitude and $x_2$ represents latitude. The coefficients a and b are used to represents the change of precipitation by longitude and latitude.



# Results{#sec-results}

All the variables can be found at @sec-model-details

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of precipitation based on longitude and latitude"
#| warning: false
#| include: false

modeldata <- data %>% group_by(longitude, latitude) %>% summarize(year_rainfall = sum(Total_Rainfall))
lm(year_rainfall ~ longitude + latitude, modeldata)
```

\begin{align} 
y = -389.9x_1 +934.7x_2 -71080.4
\end{align}


# Discussion{#sec-discussion}

## Random or regular {#sec-first-point}

When we talk about rainfall, one purpose is that the rain is formed only by natural factors. It means that the precipitation every day is totally random, just like the data we simulated in (@fig-simulate). From our daily life, if today is rainy, the precipitation is greater than not rainy day. The data we simulated shows that the rainy days are chosen randomly. We are not able to find any rules from it. However, in the real data which comes from @opendatatoronto (@fig-daily), it is totally wrong. Although it is the same that some days are rainy and some days are sunny, July seems to have more rainy days and also the precipitation is higher than other months. To find out more, the bar chart in (@fig-monthly) merges all days into months. So we can see the total rainfall for each month. We can easily see that July is higher than other weeks.
Although the rainfall seems to be random every day, the monthly data shows that the rainfall during summer is higher than other seasons.

## Longitude and latitude{#sec-second-point}

The regression model (@sec-results) shows that there are positive correlation between precipitation and longitude. Also,there are positive correlation between precipitation and latitude. This finding means that the north-eastern part of Toronto has a higher precipitation.  By analzeing this finding, the rainfall in the Toronto can be well known. It helps city workers with there work. 

## Conclusion{#sec-third-point}

In the first part, the monthly data (@fig-monthly) convinces that precipitation in the Toronto is not random every day. The highly precipitation in July concludes that summer may cause heavier rainfall than other seasons. This may cause by higher temperature in summer. In the second part, the linear regression model proves that North Eastern part of Toronto is more likely to have more precipitation. This may cause by the distance away from Ontario lake.

## Weaknesses and next steps{#sec-weeknesses}

Our data from @opendatatoronto only represent the rainfall data from January 2024 to September 2024. As a result, the change of climate by global warning may cause a change on our analysis. @rain Also, the location is only focused on Toronto. As a result, it may not suitable for all the areas around the world. Next step, the data from past years and different area can be gathered which can support or generalize our assumptions.

\newpage

\appendix

# Appendix {-}

# Additional data details{#sec-additional}
Data source: Rain Gauge Locations and Precipitation from @opendatatoronto.

Data variables: 

  date: date of the data (e.g. 2024-1-1)
  
  longitude:longitude of the rainfall
  
  latitude:latitude of the rainfall
  
  Total_Rainfall: Total rainfall that day

# Model details {#sec-model-details}
$y$: precipitation in mm

$x_1$: longitude

$x_2$: latitude


\newpage


# References


